### 2019-01-25 ex/ex-research.Rmd
add closing blurb

---
title: "Iris Research"
author: "Cole Arendt"
date: "1/6/2019"
output: pdf_document
params:
  input_data: "iris.rds"
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)

input_data <- readRDS(params$input_data)

library(dplyr)
library(tidyr)
library(ggplot2)
```

## Iris Data

This dataset is quite an interesting one. It has `r ncol(input_data)` columns
and `r nrow(input_data)` rows. We can look at a plot of the data to get a better
understanding of it:

```{r plot}
ggplot(input_data, aes(Sepal.Length, Sepal.Width)) +
  geom_point() +
  facet_grid(~Species)
```

Further, we can dig into the various groups of the data to see more detail:

```{r explore_no_output, include=FALSE}
mean_detail <- input_data %>%
  group_by(Species) %>%
  summarize_all(mean) %>%
  rename_at(vars(-Species), function(x){paste0("mean_",x)})
median_detail <- input_data %>%
  group_by(Species) %>%
  summarize_all(median) %>%
  rename_at(vars(-Species), function(x){paste0("median_",x)})

all_detail <- mean_detail %>%
  left_join(median_detail, by = c("Species"))

all_detail_tall <- all_detail %>%
  gather(key = "metric", value = "value", -Species) %>%
  spread(key = "Species", value = "value")
```

```{r explore_output}
by_med_sepal_length <- all_detail %>% arrange(desc(median_Sepal.Length)) %>% pull(Species)

knitr::kable(all_detail_tall)
```

Did you notice that `r by_med_sepal_length %>% .[[1]] %>% as.character()` has
the highest median Sepal.Length? And `r by_med_sepal_length %>%
.[[length(by_med_sepal_length)]] %>% as.character()` has the lowest.